Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-12-01 inst/paper/main.tex
2

\mypara{Intercession.} A common use case for \eval is to be combined
with \c{match.call}. What \c{match.call} does is step up the call
stack and capture the invocation of the currently executing function
and return it as an unevaluated expression. The general pattern here
is that a call to some function \c{f} is transformed to a call to
another function \c{g} with some arguments retained and others
modified. An illustration taken from the \c{vcpart} package shows a
function \c{tvcglm} that is translated to the \c{tvcm} function but
with two modifications to the argument list, argument \c{control}
can't be missing and argument \c{fit} is forced to have value \c{glm}.
The function ends with a call to \c{eval.parent} to ensure that the
rewritten call is evaluted in the same environment the original call
was.
\begin{lstlisting}
 tvcglm <- function(formula, data, family, control=tvcglm_control(), ...) {
   Call <- match.call()
   Call[[1L]] <- as.name("tvcm")
   if (!"control" %in% names(Call)) Call$control<-formals(tvcglm)$control
   if ("fit" %in% names(list(...))) warning("'fit' is ignored. ")
   Call$fit <- "glm"
   return(eval.parent(Call))
 }
\end{lstlisting}
This pattern is recognizable by the fact that the expression is a call
and the target environment is that of the parent.

\mypara{Code Generation.} The more traditional use of \eval is to
execute code that was assembled by the programmer into a string. Here
we show the method \c{plot} for class \c{sback} in package
\c{wsbackfit}, simplified for explanatory purposes. The function takes
a long argument list, the names of which are captured in the list
\c{opt}. The string \c{stub} is composed of a subset of the arguments
passed to this function; the variable is used to construct a call to
\c{base::plot} which will draw a plot. \NOTE{This example is a bit
  messed up. What is var? Does the use of ... in the eval grab the
  arguments of this call? If yes does it not end up with xlab twice?
  HELP} The \parse and \eval combination is used by the the \source
function in the base library to load R code from a file in the current
workspace.
\begin{lstlisting}
 plot.sback <- function(x,...) {
   opt <- names(list(...))
   stub <- paste(
    ifelse("xlab" %in% opt,"",paste(",xlab=\"",var,"\"",sep="")),
    ifelse("main" %in% opt,"",main.aux),
    ifelse("type" %in% opt,"",",type=\"l\""),
    sep = "")
   plot <- paste("plot(x.data,",stub,",...)",sep="")
   eval(parse(text=plot))
   ...
\end{lstlisting}
This pattern is recognized by the use of \c{parse} to turn
a string into an expression.

\mypara{Sandbox.} The combination of \eval and first class
environments can be used to construct a sandbox for evaluating
side-effecting expressions. The newly constructed environment can use
the current environment for its lexical scope. This provides the
evaluated expression access to all the bindings in the current scope
but restricts the side-effect to the newly constructed throwaway
environment. However, this sandboxing is limited because the
expression can reflectively access any environment in the call stack
or use the super assignment operator(\c{<<-}) to write to parent
scope. The \c{local} function which is part of Core R is an example of
this usage pattern.

>>>>

\mypara{Intercession.} A common use case for \eval is to be combined
with \c{match.call}. \c{match.call} walks up the call stack, captures
the code that invoked the currently executing function, and returns it
as an unevaluated expression. The pattern is to transform a call to
some function \c{f} into a call to \c{g} with some arguments retained
and others modified. As an illustration, consider the \c{vcpart}
package's function \c{tvcglm} that is translated to a call to \c{tvcm}
with two modifications to the argument list: argument \c{control}
can't be missing and \c{fit} is set  \c{"glm"}. The
function ends with a call to \c{eval.parent} to ensure that the
rewritten call is evaluted in the same environment the original call
was.
\begin{lstlisting}
 tvcglm <- function(formula, data, family, control=tvcglm_control(), ...) {
   Call <- match.call()
   Call[[1L]] <- as.name("tvcm")
   if (!"control" %in% names(Call)) Call$control<-formals(tvcglm)$control
   if ("fit" %in% names(list(...))) warning("'fit' is ignored. ")
   Call$fit <- "glm"
   return(eval.parent(Call))
 }
\end{lstlisting}
This pattern is recognizable by the fact that the expression is a call
and the target environment is that of the parent.

\mypara{Code Generation.} The more traditional use of \eval is to
execute code that was assembled by the programmer into a string. Here
we show the method \c{plot} for class \c{sback} in package
\c{wsbackfit}, simplified for explanatory purposes. The function takes
a long argument list, the names of which are captured in the list
\c{opt}. The string \c{stub} is composed of a subset of the arguments
passed to this function; the variable is used to construct a call to
\c{base::plot} which will draw a plot. \NOTE{This example is a bit
  messed up. What is var? Does the use of ... in the eval grab the
  arguments of this call? If yes does it not end up with xlab twice?
  HELP} The \parse and \eval combination is used by the the \source
function in the base library to load R code from a file in the current
workspace.
\begin{lstlisting}
 plot.sback <- function(x,...) {
   opt <- names(list(...))
   stub <- paste(
    ifelse("xlab" %in% opt,"",paste(",xlab=\"",var,"\"",sep="")),
    ifelse("main" %in% opt,"",main.aux),
    ifelse("type" %in% opt,"",",type=\"l\""),
    sep = "")
   plot <- paste("plot(x.data,",stub,",...)",sep="")
   eval(parse(text=plot))
   ...
\end{lstlisting}
This pattern is recognizable by its use of \c{parse} to turn a string
into an expression.
